---
title: "Figure 6c code"
output: html_document
---

## Source

Extracted from `C:/Users/37962/Desktop/code.Rmd`.

Related original sections: `DEG`, `UpSet`, and `Venn`.

Note: the provided Rmd contains code for the DEG table, UpSet plot, and Venn plot.
The separate left-side up/down DEG-count bar chart in panel 6c is not present as
a standalone plotting block in the provided source file; its counts come from the
same DEG table.

## Setup

```{r}
suppressPackageStartupMessages({
  library(edgeR)
  library(dplyr)
  library(tidyr)
  library(readr)
  library(ggplot2)
  library(UpSetR)
  library(ggVennDiagram)
})

strains <- c("Halomonas", "Psychrobacter", "Aequorivita", "Sphingopyxis")
strain_cols <- c(
  Halomonas="#E69F00",
  Psychrobacter="#0072B2",
  Aequorivita="#009E73",
  Sphingopyxis="#CC79A7"
)

rt <- function(x) read.delim(x, check.names=FALSE)
wt <- function(x, y) write.table(x, y, sep="\t", quote=FALSE, row.names=FALSE)
```

## DEG table used by Figure 6c

```{r}
count_files <- c(Halomonas="", Psychrobacter="", Aequorivita="", Sphingopyxis="")
sample_file <- ""
out_all <- ""
out_deg <- ""

sample_info <- rt(sample_file)
sample_info <- sample_info[!grepl("AE5500D1\\.[1-4]$", sample_info$sample), ]

run_deg <- function(strain_name, count_file) {
  x <- rt(count_file)
  rownames(x) <- x[[1]]
  x <- x[, -1, drop=FALSE]
  si <- sample_info[sample_info$strain == strain_name & sample_info$sample %in% colnames(x), ]
  si <- si[match(intersect(colnames(x), si$sample), si$sample), ]
  x <- round(as.matrix(x[, si$sample, drop=FALSE]))
  si$group <- factor(si$condition, levels=c("0B", "0D", "55B", "55D"))
  y <- DGEList(x)
  keep <- filterByExpr(y, group=si$group)
  y <- calcNormFactors(y[keep, , keep.lib.sizes=FALSE])
  design <- model.matrix(~0 + group, si)
  colnames(design) <- c("B0", "D0", "B55", "D55")
  y <- estimateDisp(y, design)
  fit <- glmQLFit(y, design)
  contrast_list <- list(
    D55_vs_D0=makeContrasts(D55-D0, levels=design),
    D55_vs_B55=makeContrasts(D55-B55, levels=design)
  )
  bind_rows(lapply(names(contrast_list), function(nm) {
    tab <- topTags(glmQLFTest(fit, contrast=contrast_list[[nm]]), n=Inf)$table
    tab$orthofinder_orthogroup <- rownames(tab)
    tab$strain <- strain_name
    tab$contrast <- nm
    tab$is_DEG <- tab$FDR < 0.05 & abs(tab$logFC) > 1
    tab$direction <- ifelse(
      tab$is_DEG & tab$logFC > 0,
      "up",
      ifelse(tab$is_DEG & tab$logFC < 0, "down", "none")
    )
    tab
  }))
}

all_deg <- bind_rows(Map(run_deg, names(count_files), count_files))
sig_deg <- all_deg %>% filter(is_DEG)
wt(all_deg, out_all)
wt(sig_deg, out_deg)
```

## UpSet plot for Figure 6c

```{r}
deg_file <- ""
out_dir <- ""

deg <- rt(deg_file) %>% filter(FDR < 0.05, abs(logFC) > 1)
if (!"DEG_direction" %in% names(deg)) deg$DEG_direction <- ifelse(deg$logFC > 0, "up", "down")

for (ct in unique(deg$contrast)) {
  dd <- deg[deg$contrast == ct, ]
  x <- split(dd$orthofinder_orthogroup, dd$strain)
  pdf(file.path(out_dir, paste0("upset_", ct, ".pdf")), width=8, height=5)
  upset(
    fromList(x[strains]),
    order.by="freq",
    sets=strains,
    mainbar.y.label="DEG count",
    sets.x.label="DEG count",
    sets.bar.color=strain_cols[strains]
  )
  dev.off()
}
```

## Venn inset for Figure 6c

```{r}
deg_file <- ""
out_dir <- ""

deg <- rt(deg_file) %>% filter(FDR < 0.05, abs(logFC) > 1)
if (!"DEG_direction" %in% names(deg)) deg$DEG_direction <- ifelse(deg$logFC > 0, "up", "down")

for (ct in unique(deg$contrast)) {
  for (di in c("total", "up", "down")) {
    dd <- deg[deg$contrast == ct, ]
    if (di != "total") dd <- dd[dd$DEG_direction == di, ]
    x <- split(dd$orthofinder_orthogroup, dd$strain)
    p <- ggVennDiagram(x[strains], label_alpha=0) + theme_void()
    ggsave(file.path(out_dir, paste0("venn_", ct, "_", di, ".pdf")), p, width=6, height=5)
  }
}
```
